---
title: "Covid19"
author: "Mikhail Samuel"
date: "2024-03-13"
output: html_document
---

## About Dataset

## Context

### Coronavirus disease (COVID-19) is an infectious disease caused by a newly discovered coronavirus. Most people infected with COVID-19 virus will experience mild to moderate respiratory illness and recover without requiring special treatment. Older people, and those with underlying medical problems like cardiovascular disease, diabetes, chronic respiratory disease, and cancer are more likely to develop serious illness.

### During the entire course of the pandemic, one of the main problems that healthcare providers have faced is the shortage of medical resources and a proper plan to efficiently distribute them. In these tough times, being able to predict what kind of resource an individual might require at the time of being tested positive or even before that will be of immense help to the authorities as they would be able to procure and arrange for the resources necessary to save the life of that patient.

### The main goal of this project is to build a machine learning model that, given a Covid-19 patient's current symptom, status, and medical history, will predict whether the patient is in high risk or not.

## Content

### The dataset was provided by the Mexican government (link). This dataset contains an enormous number of anonymized patient-related information including pre-conditions. The raw dataset consists of 21 unique features and 1,048,576 unique patients. In the Boolean features, 1 means "yes" and 2 means "no". values as 97 and 99 are missing data.

## Notes on Specific Variables

sex: 1 for female and 2 for male.

age: of the patient.

classification: covid test findings. Values 1-3 mean that the patient was diagnosed with covid in different degrees. 4 or higher means that the patient is not a carrier of covid or that the test is inconclusive.

patient type: type of care the patient received in the unit. 1 for returned home and 2 for hospitalization.

pneumonia: whether the patient already have air sacs inflammation or not.

pregnancy: whether the patient is pregnant or not.

diabetes: whether the patient has diabetes or not.

copd: Indicates whether the patient has Chronic obstructive pulmonary disease or not.

asthma: whether the patient has asthma or not.

inmsupr: whether the patient is immunosuppressed or not.

hypertension: whether the patient has hypertension or not.

cardiovascular: whether the patient has heart or blood vessels related disease.

renal chronic: whether the patient has chronic renal disease or not.

other disease: whether the patient has other disease or not.

obesity: whether the patient is obese or not.

tobacco: whether the patient is a tobacco user.

usmr: Indicates whether the patient treated medical units of the first, second or third level.

medical unit: type of institution of the National Health System that provided the care.

intubed: whether the patient was connected to the ventilator.

icu: Indicates whether the patient had been admitted to an Intensive Care Unit.

date died: If the patient died indicate the date of death, and 9999-99-99 otherwise.

```{r}
library(tidyverse)
library(tidymodels)
library(caret)
library(plotly)
library(skimr)
library(rpart)
library(e1071)
library(randomForest)
```

```{r}
covid19  <- read_csv("/Users/yn/mikhail/data science/jupeter/workspace 8/Covid Data.csv",show_col_types = FALSE)
head(covid19)
```

```{r}
skim(covid19)
```

### Most people in 25 - 55 years old

```{r}
sex <- table(covid19$SEX)
names(sex)  <- c('female','male')
sex

p <- qplot(data = covid19,x = AGE,geom = 'histogram') + 
facet_wrap(~SEX,labeller = labeller(SEX = c('1' = 'female','male')))

ggplotly(p)
```

### Most people classified as in 1-3 sepcificly in 3 it means they have covid19

```{r}
p <- covid19 %>%
ggplot(aes(AGE,fill = as.factor(CLASIFFICATION_FINAL))) +
geom_histogram()+
labs(
    fill = 'CLASIFFICATION_FINAL'
)

ggplotly(p)
```

### Most of them in home not hospital it means tow things:

#### `1- They don't dengrous conditions 'I wich to be'`

#### `2- Hospitals have'nt beds to them`

```{r}
p <- covid19 %>%
ggplot(aes(AGE,fill = as.factor(PATIENT_TYPE))) +
geom_histogram()+
labs(
    fill = 'PATIENT_TYPE'
)

ggplotly(p)
```

### Most of them haven't PNEUMONIA

```{r}
p <- covid19 %>%
ggplot(aes(AGE,fill = as.factor(PNEUMONIA))) +
geom_histogram(position = 'dodge',binwidth = 6) +
scale_x_continuous(limits = c(0,80))+
labs(
    fill = 'PNEUMONIA'
) 

ggplotly(p,dynamicTicks = TRUE)
```

### Most of females not PREGNANT

```{r}
p <- covid19 %>% filter(SEX == 1) %>% 
ggplot(aes(AGE,fill = as.factor(PREGNANT))) +
geom_histogram()+
labs(
    fill = 'PREGNANCY'
)

ggplotly(p)
```

### Most of them haven't DIABETES

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(DIABETES))) +
geom_histogram(position = 'dodge',binwidth = 6)+
scale_x_continuous(limits = c(0,80)) +
labs(
    fill = 'DIABETES'
)

ggplotly(p)
```

### It very Mostly all of them haven't `(COPD)` Chronic obstructive pulmonary disease

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(COPD))) +
geom_histogram()+
labs(
    fill = 'COPD'
)

ggplotly(p)
```

### Most of theem haven't ASTHMA

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(ASTHMA))) +
geom_histogram()+
labs(
    fill = 'ASTHMA'
)

ggplotly(p)
```

### All of them haven't INMSUPR

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(INMSUPR))) +
geom_histogram()+
labs(
    fill = 'INMSUPR'
)

ggplotly(p)
```

### Most of them haven't HIPERTENSION

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(HIPERTENSION))) +
geom_histogram()+
labs(
    fill = 'HIPERTENSION'
)

ggplotly(p)
```

### Most of them haven't OTHER_DISEASE

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(OTHER_DISEASE))) +
geom_histogram()+
labs(
    fill = 'OTHER_DISEASE'
)

ggplotly(p)
```

### Most of them haven't CARDIOVASCULAR

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(CARDIOVASCULAR))) +
geom_histogram()+
labs(
    fill = 'CARDIOVASCULAR'
)

ggplotly(p)
```

### Most of them haven't OBESITY

```{r}
p <- covid19 %>% ggplot(aes(AGE,fill = as.factor(OBESITY))) + geom_histogram(position = 'dodge',binwidth = 6)+ scale_x_continuous(limits = c(0,80)) + labs( fill = 'OBESITY' )

ggplotly(p,dynamicTicks = TRUE)
```

### Most of them haven't RENAL_CHRONIC

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(RENAL_CHRONIC))) +
geom_histogram()+
labs(
    fill = 'RENAL_CHRONIC'
)

ggplotly(p)
```

### Most of them don't smoke TOBACCO

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(TOBACCO))) +
geom_histogram()+
labs(
    fill = 'TOBACCO'
)

ggplotly(p)
```

### Most of them don't admitte in ICU

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(ICU))) +
geom_histogram(position = 'dodge',binwidth = 6)+
scale_x_continuous(limits = c(0,80)) +
labs(
    fill = 'ICU'
)

ggplotly(p)
```

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(USMER))) +
geom_histogram()+
labs(
    fill = 'USMER'
)

ggplotly(p)
```

```{r}
p <- covid19 %>% 
ggplot(aes(AGE,fill = as.factor(MEDICAL_UNIT))) +
geom_histogram()+
labs(
    fill = 'MEDICAL_UNIT'
)

ggplotly(p)
```

## Handle Col's to predict

```{r}
covid19II <- covid19  %>% select(-DATE_DIED,-SEX,-AGE)
head(covid19II)

covid19II<- as.data.frame(map(
  covid19II,~replace(
    .,
    which(. %in% c(97,98,99)),
    0
    )
  ))

head(covid19II)
```

## Create Train and Test set

```{r}
set.seed(19)
index <- createDataPartition(covid19II$CLASIFFICATION_FINAL,p = .8,list = FALSE)
trainSet <- covid19II[index,]
testSet <- covid19II[-index,]
```

## Fit and Predict Model

```{r}
fit <- glm(CLASIFFICATION_FINAL ~ . ,data = trainSet)

fit
```

```{r}
#predict
pre <- predict(fit,newdata = testSet)
# plot predict with truth
plot(testSet$CLASIFFICATION_FINAL,pre)
# I'll put predict in testSet to predict be the same levels in the same data
testSet <- testSet |> mutate(pre =pre)

# RMSE
rmse(data = testSet,truth = CLASIFFICATION_FINAL,estimate = pre)
# R-Squard
rsq(data = testSet,truth = CLASIFFICATION_FINAL,estimate = pre)

# Plot fit
plot(fit)
```

```{r}
fit <- naiveBayes(CLASIFFICATION_FINAL ~ . ,data = trainSet)

fit

```

```{r}
#predict
pre <- predict(fit,newdata = testSet)
# plot predict with truth
plot(testSet$CLASIFFICATION_FINAL,pre)
# I'll put predict in testSet to predict be the same levels in the same data
testSet <- testSet |> mutate(pre =pre)

# RMSE
rmse(data = testSet,truth = CLASIFFICATION_FINAL,estimate = pre)
# R-Squard
rsq(data = testSet,truth = CLASIFFICATION_FINAL,estimate = pre)

```

```{r}
newcovid19II<- as.data.frame(map(
  covid19,~replace(
    .,
    which(. %in% c(97,98,99)),
    0
    )
  ))

write_csv(newcovid19II,'covid19II.csv')
```
